Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-07-13 paper/manuscript.tex
Updates from Overleaf

\subsection{Absolute and relative calculations must deal with some of the same issues}
\subsubsection{Handling weak binders and high dissociation rates}
\todo[inline, color={green!20}]{ASJM: This a bit waffly. What are the recommendations, of how to setup these restraints?}
\todo[inline, color={red!40}]{AR: Took a pass at this.}
In binding free energy calculations, only the conformations in which the receptor and ligand form a bound complex should be sampled from the bound states (Section~\ref{sec:theory}).
In absolute calculations, this applies to the fully interacting state in the complex leg of the thermodynamic cycle (top-right state in Figure~\ref{fig:absolute-thermodynamic-cycle}), while in relative calculations this must be true at both end states of the complex leg (top- and bottom-right states in Figure~\ref{fig:binding-thermodynamic-cycle}).
In principle, this requires defining which conformations are considered to be bound before running the calculation, but it is common practice to start the simulation with the ligand already placed in the binding site and rely on kinetic trapping to maintain the bound complex.
This strategy, however, can fail when the dissociation rate of the ligand has the same or smaller order of magnitude than the length of the simulation.
This is typical of weak binders such as fragments binding shallow pockets with $\mu$M-mM affinities~\cite{georgiou_pushing_2017,pan2017quantitative}.
In this case, using a flat-bottom or harmonic restraint between receptor and ligand in the bound state(s) can prevent dissociations~\cite{georgiou_pushing_2017,rizzi2019sampl6}.
We stress that this is normally avoided as it generally introduces bias in the free energy estimate, which is why the restraint is usually activated only in the intermediate states in absolute calculations.
The bias can be corrected through reweighting schemes~\cite{rizzi2019sampl6}, but this post-processing step can be avoided if a flat-bottom restraint is used and the ligand never hits the potential wall during the simulation in the bound state.
It is important to note that the spring constant and/or radius parameters of the restraint effectively determine which conformations are considered to be bound.
As a consequence, these parameters must be tuned to the system so that only the binding site is accessible to the ligand.
Again, this step is particularly important for weak binders as their free energy of binding is known to be more sensitive to the definition of the binding site~\cite{gilson1997statisticalthermodynamic}.

@@

%
\paragraph{Reversible binding simulations}
An even more stringent test of the correctness of binding free energy calculations is to compare the results to the equilibrium binding constants derived from long timescale reversible binding simulations~\cite{pan2017quantitative}.  For small ligands with millimolar affinities, repeated binding to and unbinding from the protein can occur for a large number of times in a sufficiently long unbiased MD simulation (10-100$\mu$s), and the equilibrium binding constants can be straightforwardly computed from the ratio of bound to unbound fractions of the simulation time. The agreement between the binding free energy calculations and the reversible binding simulations--given the same system preparation and the same force field parameters--will strongly support the correctness of both calculations, as the same results are arrived at by two independent methods, and any discrepancy will suggest some systematic error in one, or both, of the two methods. As part of validation testing of alchemical free energy codes a benchmark set to compare alchemical and direct computation of equilibrium binding constant should become standard in future.
%
\subsection{Recommendations}
\todo[inline]{ASJSM:Maybe here we have a quick summary on what things should be checked for to make sure simulations are ok?}
\begin{itemize}
\item Examining output data for common problems with discussions of what exactly to plot or look at; examples of typical curves for dV/dlambda and free energy versus lambda, for example
\begin{itemize}
\item Make sure ligand sampling in the binding site is reasonable for it's expected potency and fit, and that it does not depart out of binding site. 
\item Significant discrepancies between free energies calculated with different free energy estimators (TI, BAR, MBAR)
\item Poor replica mixing (for replica-exchange) or lambda-space sampling for single-replica methods
\item Correlation time as a function of lambda that does not vary relatively smoothly
\item Dependence of the free energy on initial conformation
\item Torsions that have multiple low-energy minima that are visited rarely or not at all. Also, comparing torsional behavior of ligands in solvent and complex simulations 
\item The convergence of the free energy change as a function of lambda, or of simulation time (Figure 11).
\item Diagnostics for non-equilibrium methods, robustness for switching time etc.
\item More “usual suspects”
\end{itemize}
\item Other considerations for many transformations
\end{itemize}
%
\subsection{Best practices for reporting data }
\label{sec:plot_data}
Data generation and adequate analysis is not sufficient to finish a study. As a practitioner of alchemical free energy simulations you should use best practice for reporting and plotting your results. We  encourage the following standard set of analyses and ways to represent data. 
\paragraph{Statistics to include}
As with any modelling technique, misuse of statistical analysis can skew the perception of how well models perform in free energy predictions. First, error estimates should always be included on your predictions in whatever form you present your data (scatterplots, barplots, etc; see next paragraph). We recommend performing triplicates of your predictions at minimum to ensure some measure of reliability in your data. It may seem excessive, but  uncertainty estimates often underestimate the true statistical uncertainty. Where performing multiple replicas of the simulation is not possible, lower error bounds such as given by the MBAR estimator should be included.

Because of the strong drug discovery context and the relevance of alchemical free energy predictions to structure activity relationship (SAR) projects, both the overall correlation of your predictions relative to experiment and the ability of your models to detect the best-scoring compounds (i.e. rankings) should be computed. Conventionally, this means including an R\textsuperscript{2} (or Pearson's R: +1 high correlation 0 no correlation -1 high anti-correlation) and a Kendall \texttau{} (with perfect ranking agreement when \texttau=1 and perfect disagreement when \texttau=-1) metric in your results. Additionally, practitioners may choose to include a Spearman \textrho{} as well. Brown et al.~\cite{brown2009healthy} have provided a useful analysis in terms of upper bounds of expected possible correlations between experiment and computation with a given potency range for the compounds. For example, for potency ranges of 2 log units it would be impossible to get a higher correlation in R than 0.8 because of experimental uncertainties~\cite{brown2009healthy}. What often is neglected to include is an error analysis on correlation statistics that arise from the errors of both experimental and computed data. One way to include such error analysis for correlation metrics is using bootstrapping on the datasets. The D3R community challenges follows best practices on their data evaluation with readily available python scripts online~\cite{2018drugdata}, based on work by Pat Walters~\cite{walters2013what}. Other analysis software also provide similar functionality for bootstrapping datasets~\cite{antonia2019michellab}. 
%
Mean unsigned error (MUE, also called mean absolute error/MAE) is another key statistic to include in your results. Even though some models' near-perfect correlation and ranking statistics might suggest excellent accuracy, MUE values can still show multiple kcal/mol of error, providing important additional insight into performance. Furthermore, MUE allows for unbiased comparisons between predictive models as it is less sensitive to dataset size. Other metrics such as Gaussian Random Affinity Model (GRAM)~\cite{cui2019gram}, Predictive Interval (PI) Relative Absolute Error (RAE), that attempt to correct for the inherent potency range of a dataset, which can aid in comparing success between different targets. Even though this section may serve as an introduction to the main statistical practices, the authors recommend further reading on evaluation of computational models \cite{jain2008recommendations, walters2013what, brown2009healthy, walterthoughts}.

Reporting the results of relative free energy calculations require an additional level of care. As shown in Figure 5, relative free energies can be performed in two arbitrary directions, thus relative free energies may be reported as either positively or negatively valued, depending on the choice of what is forward. The consequence of the two possible signs for relative free energies is that correlation statistics (such as Pearson's R and Kendall \texttau{}) can be skewed depending on which sign is analysed. The issue of this inconsistency can be circumvented by either plotting all datapoints within a consistent quadrant\cite{perez-benito2019predicting}, or by avoiding the use of correlation statistics for assessment of relative free energy calculations and instead measuring accuracy using RMSE and MUE which are unaffected by choice of sign.
%

\paragraph{Presenting your data}
As essentially all alchemical free energy prediction schemes are regression problems, the preferred type of plot is a scatter plot (see Figure \ref{fig:scatterplot_analysis}). Most alchemical FE projects will concern 10-50 ligands; any study with \textless10 ligands is more suitable for bar plots (with inclusion of error bars). Any study with \textgreater50 ligands should be trellised to not overestimate statistics. In that respect, it is bad practice to place multiple datasets on the same plot because this can suggest high model accuracy even though the individual models perform less well \cite{walterthoughts}.
%

Because we are interested mainly in the linear relationship between the alchemical FE predictions and the experimentally-determined affinity values,  and if units are consistent for both axis, plots should be depicted with the same range on both axes (i.e. $x=y$) with a 1:1 aspect ratio. Furthermore, bounds should be depicted for the 1- and 2-kcal/mol confidence regions. These regions can serve as tools to communicate your model performance: any predictions inside the 1 kcal/mol region can be seen as highly reliable, any predictions inside the 2 kcal/mol region should be seen as somewhat reliable, and any predictions outside the confidence regions should be expected to be unreliable (and handled as outliers). In a drug discovery context, this type of data depiction may suggest the reliability of alchemical FE predictions in the project, and can give an idea of how trustworthy predictions can be for synthesis ideas. 

>>>>

\subsection{Absolute and relative calculations deal with some of the same issues}
\subsubsection{Handling weak binders and high dissociation rates}
\todo[inline, color={green!20}]{ASJM: This a bit waffly. What are the recommendations, of how to setup these restraints?}
\todo[inline, color={red!40}]{AR: Took a pass at this.}
In binding free energy calculations, only the conformations in which the receptor and ligand form a bound complex should be sampled from the bound states (Section~\ref{sec:theory}).
In absolute calculations, this applies to the fully interacting state in the complex leg of the thermodynamic cycle (top-right state in Figure~\ref{fig:absolute-thermodynamic-cycle}), while in relative calculations this must be true at both end states of the complex leg (top- and bottom-right states in Figure~\ref{fig:binding-thermodynamic-cycle}).
In principle, this requires defining which conformations are considered to be bound before running the calculation, but it is common practice to start the simulation with the ligand already placed in the binding site and rely on kinetic trapping to maintain the bound complex.
This strategy, however, can fail when the dissociation rate of the ligand has the same or smaller order of magnitude than the length of the simulation.
This is typical of weak binders such as fragments binding shallow pockets with $\mu$M-mM affinities~\cite{georgiou_pushing_2017,pan2017quantitative}.
In this case, using a flat-bottom or harmonic restraint between receptor and ligand in the bound state(s) can prevent dissociations~\cite{georgiou_pushing_2017,rizzi2019sampl6}.
We stress that this is normally avoided as it generally introduces bias in the free energy estimate, which is why the restraint is usually activated only in the intermediate states in absolute calculations.
The bias can be corrected through reweighting schemes~\cite{rizzi2019sampl6}, but this post-processing step can be avoided if a flat-bottom restraint is used and the ligand never hits the potential wall during the simulation in the bound state.
It is important to note that the spring constant and/or radius parameters of the restraint effectively determine which conformations are considered to be bound.
As a consequence, these parameters must be tuned to the system so that only the binding site is accessible to the ligand.
Again, this step is particularly important for weak binders as their free energy of binding is known to be more sensitive to the definition of the binding site~\cite{gilson1997statisticalthermodynamic}.

@@

%
\paragraph{Reversible binding simulations}
An even more stringent test of the correctness of binding free energy calculations is to compare the results to the equilibrium binding constants derived from long timescale reversible binding simulations~\cite{pan2017quantitative}.  For small ligands with millimolar affinities, repeated binding to and unbinding from the protein can occur for a large number of times in a sufficiently long unbiased MD simulation (10-100$\mu$s), and the equilibrium binding constants can be straightforwardly computed from the ratio of bound to unbound fractions of the simulation time. The agreement between the binding free energy calculations and the reversible binding simulations--given the same system preparation and the same force field parameters--will strongly support the correctness of both calculations, as the same results are arrived at by two independent methods, and any discrepancy will suggest some systematic error in one, or both, of the two methods. As part of validation testing of alchemical free energy codes a benchmark set to compare alchemical and direct computation of equilibrium binding constant should become standard in future.
%
\subsection{Recommendations}
\todo[inline]{ASJSM:Maybe here we have a quick summary on what things should be checked for to make sure simulations are ok?}

\item It is important to carefully examine output data for common problems.  Some of the most important things to check for rae:
\begin{itemize}
\item Make sure the ligand samples the binding site reasonably tightly for it's expected potency and fit, and that it does not depart out of binding site in the coupled end state if it is a moderate to strong binder. 
\item Significant discrepancies (further outside the error estimates than would be plausible) between free energies calculated with different free energy estimators such as TI, BAR, and MBAR. All of these estimators converge to the same results with sufficient sampling.  Differences between them indicate poor overlap or errors in processing.
\item Poor replica mixing (for replica-exchange) or lambda-space sampling for single-replica methods. If the system is not mixing between states, then  the states are insufficiently close for mixing, or else there are bottlenecks in the configurational sampling that limit the accuracy.
\item Correlation time  that does not vary relatively smoothly as a function of $\lambda$.  Discontinuities indicate that the system is sampling significantly different configurations when the Hamiltonian changes only a bit, which usually indicates sampling problems.
\item Dependence of the free energy on initial conformation of the system. Clearly, ensemble average properties should not depend on the starting point.
\item Torsions with multiple low-energy minima where some of these minima are that are visited rarely or not at all. Which torsions have low energy minima can best be found by comparing to the simulation in the solvent. There should be clear physical reasons that  simulation in the complex has different torsional distributions that the ligand in the solvent. 
\item The free energy should vary relatively smoothly with $\lambda$. If it varies drastically, then either there need to be finer sampling in $\lambda$ in this region, or there are sampling problems there.
\item The free energy should clearly converge as a function or of simulation time (Figure $\ref{fig:convergence_forward_reverse}$).
\item If using nonequilbirum methods, is the result independent of the speed at which the nonequilibrium change is performed?  Nonequilibrium methods are in theory independent of the switching time in the limit of good sampling unless the switching time is simply too short. 
%\item Other Diagnostics for non-equilibrium methods, robustness for switching time etc.?
\item In general, look at the output data such as energies and see if it matches your expectations.and also visualize the simulation trajectories themselves to see if it matched expectations.  There are many errors that are caught by only one of the two approaches
%\item More “usual suspects”
\end{itemize}
%\item Other considerations for many transformations

\subsection{Best practices for reporting data }
\label{sec:plot_data}
Carrying out good data generation and proper analysis is not sufficient to finish a study. As a practitioner of alchemical free energy simulations you also  should use best practices for reporting and plotting your results. We  encourage the following standard set of analyses and ways to represent data. 
\paragraph{Statistics to include}
As with any modelling technique, misuse of statistical analysis can skew the perception of how well models perform in free energy predictions. First, error estimates should always be included on your predictions in whatever form you present your data (scatterplots, barplots, etc; see next paragraph). We recommend performing triplicates of your predictions at minimum, with starting points that are expected to be uncorrelated, to ensure some measure of reliability in your data. This replication may seem excessive, but  uncertainty estimates often underestimate the true statistical uncertainty. Where performing multiple replicas of the simulation is not possible, lower error bounds such as given by the MBAR estimator should be included.

Because of the recent importance of alchemical free energy predictions in the  drug discovery context, and the relevance of these quantities to structure activity relationship (SAR), both the overall correlation of your predictions relative to experiment and the ability of your models to detect the best-scoring compounds (i.e. rankings) should be computed. Conventionally, this means including an R\textsuperscript{2} (or Pearson's R: +1 high correlation 0 no correlation -1 high anti-correlation) and a Kendall \texttau{} (with perfect ranking agreement when \texttau=1 and perfect disagreement when \texttau=-1) metric in your results. Additionally, practitioners may choose to include a Spearman \textrho{} as well. Brown et al.~\cite{brown2009healthy} have provided a useful analysis in terms of upper bounds of expected possible correlations between experiment and computation with a given potency range for the compounds. For example, for potency ranges of 2 log units it would be impossible to get a higher correlation in R than 0.8 because of experimental uncertainties~\cite{brown2009healthy}. What often is neglected to include is an error analysis on correlation statistics that arise from the errors of both experimental and computed data. One way to include such error analysis for correlation metrics is using bootstrapping on the datasets. The D3R community challenges follows best practices on their data evaluation with readily available python scripts online~\cite{2018drugdata}, based on work by Pat Walters~\cite{walters2013what}. Other analysis software also provide similar functionality for bootstrapping datasets~\cite{antonia2019michellab}. 
%
Mean unsigned error (MUE, also called mean absolute error/MAE) is another key statistic to include in your results. Even though some models' near-perfect correlation and ranking statistics might suggest excellent accuracy, MUE values can still have errors of show multiple kcal/mol of error, providing important additional insight into performance. Furthermore, MUE allows for unbiased comparisons between predictive models as it is less sensitive to dataset size. Other metrics such as Gaussian Random Affinity Model (GRAM)~\cite{cui2019gram}, Predictive Interval (PI) Relative Absolute Error (RAE), that attempt to correct for the inherent potency range of a dataset, which can aid in comparing success between different targets. We recommend further reading on evaluation of computational models \cite{jain2008recommendations, walters2013what, brown2009healthy, walterthoughts}.

Reporting the results of relative free energy calculations requires care. As shown in Figure~\ref{fig:5}, relative free energies can be performed arbitrarily as a forward or a reverse process, and thus relative free energies may be reported as either positively or negatively valued. The consequence of the two possible signs for relative free energies is that correlation statistics (such as Pearson's R and Kendall \texttau{}) can be skewed depending on which sign is analysed. The issue of this inconsistency can be circumvented by either plotting all datapoints within a consistent quadrant\cite{perez-benito2019predicting}, or by avoiding the use of correlation statistics for assessment of relative free energy calculations and instead measuring accuracy using RMSE and MUE which are unaffected by choice of sign.
%

\paragraph{Presenting your data}
As essentially all alchemical free energy prediction schemes are regression problems, the preferred type of plot is a scatter plot (see Figure \ref{fig:scatterplot_analysis}). Most alchemical FE projects will concern 10-50 ligands; any study with \textless10 ligands is more suitable for bar plots (with inclusion of error bars). Any study with \textgreater50 ligands should be trellised to not overestimate statistics. Because of this, it is bad practice to place multiple datasets on the same plot, because this can suggest high model accuracy even though the individual models perform less well \cite{walterthoughts}.
%

Because we are interested mainly in the linear relationship between the alchemical FE predictions and the experimentally-determined affinity values, plots should be depicted with the same range on both axes (i.e. $x=y$) with a 1:1 aspect ratio, with units for both experiment and simulation converted to be the same. Furthermore, bounds should be depicted for the 1- and 2-kcal/mol confidence regions. These regions can serve as tools to communicate your model performance: any predictions inside the 1 kcal/mol region can be seen as highly reliable, any predictions inside the 2 kcal/mol region should be seen as somewhat reliable, and any predictions outside the confidence regions should be expected to be unreliable (and handled as outliers). In a drug discovery context, this type of data depiction may suggest the reliability of alchemical FE predictions in the project, and can give an idea of how trustworthy predictions can be for synthesis ideas. 
